---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HamONdest

<!-- badges: start -->
<!-- badges: end -->

HamONdest is a data-package which contains a growing set of potential destinations and their estimated 'supplies' in Hamilton, Ontario, Canada. It currently contains schools, municipally-owned land, health care and health and residential care facilities.

The aim of this data-package is to facilitate the calculation of accessibility and mobility measures in the transportation planning context. The majority of data is sourced from [Hamilton's Open Data Portal](https://open.hamilton.ca/) and is augmented with additional open data from higher levels of government (provincial and federal), predicted values, and/or data from [OpenStreetMap](https://www.openstreetmap.org).

The HamONdest package is still a work in progress and is available [here](https://github.com/soukhova/HamONDest-package) and the vignettes can be viewed [here](https://soukhova.github.io/HamONdest-package/articles/HamONdest-vignette.html)


## Installation

You can install the released version of HamONdest using:

``` {r, eval = FALSE}
if(!require(remotes)){
    install.packages("remotes")
    library(remotes)
}
remotes::install_github("soukhova/HamONdest-package",
                         build_vignettes = TRUE)
```

## Examples

```{r libraries, eval=TRUE, message=FALSE, warning=FALSE}
library(HamONdest)
library(ggplot2)
library(dplyr)
library(sf)
```

The data package contains the following destinations in Hamilton, Ontario.

### Schools
Elementary and Secondary Public and Public Catholic Schools in **2011**, size based on on-the-ground-capacity (OTGC)
```{r schools-2011, eval=TRUE, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white") +
  geom_sf(data = Schools_201516_201011 %>% filter(Year != "2016"),
          aes(col = OTGC2011),
          shape = 1,
          size = 3) +
  scale_color_distiller(palette = "Spectral")
```

Elementary and Secondary Public and Public Catholic Schools in **2016**, size based on on-the-ground-capacity (OTGC)
``` {r schools-2016, eval=TRUE, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white")+
  geom_sf(data = Schools_201516_201011 %>% filter(Year != "2011"),
          aes(col = OTGC2016),
          shape = 1,
          size = 3) +
  scale_color_distiller(palette = "Spectral")
```

### Municipally-Owned Properties
Up-to-date City-Owned Property(Parks, Open Space, Vacant Land, + Properties with civil services like fire stations, police station, community centres, etc.) with person destination potential. Sourced from Open Data Hamilton.
```{r city-owned-properties, eval=TRUE, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white")+
  geom_sf(data = City_Owned_Property,
          aes(col = CATEGORY_TYPE),
          shape = 1) 
```

### Health and Community Care Service Providers
All Health and community care facilities (Clinics, Hospitals, Residential Care Facilities, Long-term Care Facilities, etc.) as sourced from City of Hamilton Open Data and Ontario GeoHub:
```{r all-care-facilities, eval=TRUE, warning=FALSE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white")+
  geom_sf(data = Care_Facilities,
          aes(col = Type),
          shape = 1) 
```

For example, locations to a specific service provider type can be viewed. Below is a plot of Hospitals:
```{r hospitals, eval=TRUE, warning=FALSE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white")+
  geom_sf(data = Care_Facilities %>% filter(Type == "Hospital"),
          color = 2,
          shape = 1)
```

## More Information 
Coordinate system of all objects contained in this data-package is as shown:
```{r coordinates, eval=TRUE}
sf::st_crs(Schools_201516_201011)
```

This data package is still a work in progress. See additional visualizations in the Vignettes. 


