---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HamONdest

<!-- badges: start -->
<!-- badges: end -->

The goal of HamONdest is to compile all potential destinations (Schools, municipally-owned land, parks, libraries, etc.) and their estimated 'demands' in Hamilton, Ontario, Canada all in one data-package to facilitate the calculation of accessibility and mobility measures in the transportation planning context. The majority of data is sourced from [Hamilton's Open Data Portal](https://open.hamilton.ca/) and is augmented with additional open data from higher levels of government (provincial and federal), predicted values, and/or data from [OpenStreetMap](https://www.openstreetmap.org).


## Installation

You can install the released version of HamONdest using:

``` {r, eval = FALSE}
if(!require(remotes)){
    install.packages("remotes")
    library(remotes)
}
remotes::install_github("soukhova/HamONdest-package")
```

## Examples

```{r libraries, eval=TRUE}
library(HamONdest)
library(ggplot2)
library(dplyr)

Schools_201516_201011 <- HamONdest::Schools_201516_201011
Schools_Catchments_201516_201011 <- HamONdest::Schools_Catchments_201516_201011
Ham_CityBound <- HamONdest::Ham_CityBound
```


The data package contains the following destinations in Hamilton, Ontario.

Elementary and Secondary Public and Public Catholic Schools in 2011, size based on on-the-ground-capacity (OTGC)
```{r, eval=TRUE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white") +
  geom_sf(data = Schools_201516_201011 %>% filter(Year != "2016"),
          aes(col = OTGC2011),
          shape = 1,
          size = 3) +
  scale_color_distiller(palette = "Spectral")
```

Elementary and Secondary Public and Public Catholic Schools in 2016, size based on on-the-ground-capacity (OTGC)
``` {r, eval=TRUE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white")+
  geom_sf(data = Schools_201516_201011 %>% filter(Year != "2011"),
          aes(col = OTGC2016),
          shape = 1,
          size = 3) +
  scale_color_distiller(palette = "Spectral")
```

The change in Elementary and Secondary Public and Public Catholic Schools between 2011 and 2016.
```{r, eval=TRUE}
ggplot() +
  geom_sf(data = Ham_CityBound,
          size = 0.5,
          alpha = 0.5,
          color  = "black",
          fill = "white")+
  geom_sf(data = Schools_201516_201011 %>% filter(Status != "NoChange"),
          aes(col = Status, size = Level),
          shape = 1) 
```

Coordinate system and projection
```{r, eval=TRUE}
sf::st_crs(Schools_201516_201011)
```

## More Information 

This data package is still a work in progress.
